---
title: "4 Years of Reporting on the Full Scale Russian Invasion of Ukraine"
output: html
---



```{r loading packages}

#install.packages('guardianapi')
#install.packages('usethis')
#install.packages('ggthemes')
#install.packages('textdata')
#install.packages('widyr')

# Tidyverse paradigm
library(dplyr)
library(tidyverse)
library(widyr)


# API and webscraping
library(rvest)
library(httr2)
library(guardianapi)


# Visualizations
library(ggplot2)
library(patchwork)
library(ggthemes)
library(ggrepel)
library(RColorBrewer)
library(scales)


# Text analyisis
library(quanteda)
library(quanteda.textstats)
library(quanteda.textplots)
library(quanteda.textmodels)
library(textstem)
library(tidytext)
library(textdata)
library(topicmodels)



```


```{r Kyiv Post Web Scraping}

# Initializing the links for scraping
kp_default_url <- 'https://www.kyivpost.com'
kp_war_url <- 'https://www.kyivpost.com/topic/war-in-ukraine'

# Function to check if anything is missing and returns missing character if so
na_if_empty <- function(x) {
  if (length(x) == 0 || all(x == '') || all(is.na(x))) { # if any kind of 0 or non-character
    return(NA_character_) # return NA character
  } else {
    return(x) 
  }
}


# Initializing a function that reads the article pages
read_article_body <- function (kp_article_links) { 
  kp_dfs <- list() # Initizalizing an empty list
  for(i in seq_along(kp_article_links)) { # for loop that loops over 


    # Closing all connections every 10 pages
    if (i %% 10 == 0) {
      closeAllConnections()
      gc(verbose = FALSE)
    }
    print(paste0('Scraping content from article ', i, ' of ', length(kp_article_links))) # Text message to see how far along the scraping is
    if (is.na(kp_article_links[i])) next # Skips over dead pages

    # Second fail safe - catches errors and returns NULL instead of crashing
      page_html <- tryCatch({
  read_html(kp_article_links[i]) 
}, error = function(e) {
  closeAllConnections() # Closes broken handle if an article fails
  return(NULL)
})
    if (is.null(page_html)) next # if null - skips over it

    # Parse the content wrappers of each url page 
    kp_content_wrapper <- page_html %>%
      html_elements('.content-wrapper')
    

    # Parse the titles out of the wrapper as text
    kp_titles <- kp_content_wrapper %>%
      html_elements('.post-title') %>% 
      html_text2() %>%
      na_if_empty()

    # Parse the authors out of the wrapper as text
    kp_content_authors <- kp_content_wrapper %>% 
      html_elements('.post-info a') %>%
      html_text2() %>%
      na_if_empty()

    # Parse the date out of the wrapper as text - needs more cleaning afterwards
    kp_content_date_uncleaned <- kp_content_wrapper %>%
      html_elements('.post-info') %>%
      html_text2() %>%
      na_if_empty()

    # Parse the lead text out of the wrapper as text
    kp_content_lead <- kp_content_wrapper %>%
      html_elements('.post-lead') %>% 
      html_text2() %>%
      na_if_empty()

    # Parse the body text out of the wrapper as text - needs more cleaning afterwards
    kp_body_text_uncleaned <- kp_content_wrapper %>%
      html_elements('#post-content p, #post-content li') %>%
      html_text2() %>%
      paste(collapse = "\n\n") %>%
      na_if_empty()
      
    # Add everything, including the links, to a tibble 
    kp_tibble <- tibble(
    title = kp_titles,
    lead_text = kp_content_lead,
    body_text_uncleaned = kp_body_text_uncleaned,
    date_uncleaned = kp_content_date_uncleaned,
    authors = kp_content_authors, 
    link = kp_article_links[i])

    # add them to the empty list
    kp_dfs[[i]] <- kp_tibble

    # Adding sleep at .5 seconds
    Sys.sleep(.1)

  }

  # Bind rows together and return them
  return(bind_rows(kp_dfs))

}



# Initializing function that reads through the pages of the website itself in the war section
read_war_page <- function(kp_war_url, max_page)   { 

all_pages_dfs <- list() # initialize empty list to put in all of the pages
for (p in 1:max_page) { # Looping across all of the pages (max page is automatically parsed)
  kp_war_page <- paste0(kp_war_url, '?page=', p) # Changing through the pages
  print(paste0('Processing page number: ', p, ' of ', max_page)) # printing out which page it is currently parsing


  # Every 10 pages, close all connections to free memory
  if (p %% 10 == 0) {
    closeAllConnections()
    gc(verbose = FALSE)
  }

  # Fail safe 
 kp_html_page_specific <- tryCatch({
  read_html(kp_war_page)
}, error = function(e) {
  closeAllConnections() # Closes broken handle if main page fails
  return(NULL)
})
  if (is.null(kp_html_page_specific)) next




  # Parsing all the elements  under the category latest
  kp_html_elements <- kp_html_page_specific %>%
  html_elements('.post-info') %>%
  na_if_empty()

  # Parsing the links
  kp_article_links <- kp_html_elements %>%
  html_elements('h2 > a') %>%
  html_attr('href', default = NA_character_) %>%
  url_absolute(kp_default_url) %>%
  na_if_empty()

# Using the previously initialized function to scrape the articles while going through the pages and adding them to the empty list
all_pages_dfs[[p]] <- read_article_body(kp_article_links)

Sys.sleep(3) # sleep function of 3 seconds
}
return(bind_rows(all_pages_dfs)) # binding the of the list together and returning them
}


# The overarching function that incorporates the previous two functions and caches 
kp_scraper <- function(cache_file = 'kp_articles.rds', force_refresh = FALSE) { # Caches the file 
  if (file.exists(cache_file) && !force_refresh) { # Checks if the file exists in cache and not force refreshed
  file_age <- difftime(Sys.time(), file.mtime(cache_file), 
  units = 'weeks') # If the file already exists checks how old it is in weeks
  if (file_age < 4) { # If file age is less than 4 weeks
    print(paste0('loading cached data - ', round(file_age, 1), 'week(s) old')) # Returns message that it loads the cached file and mentions age in weeks
    return(readRDS(cache_file)) # Returns the cached file
  }
}

# If the file is older than 4 weeks and the function isn't force_refreshed 
# Automatically checks the max pages
max_page <- read_html(kp_war_url) %>% 
  html_element('li.page-item:nth-child(6) > a:nth-child(1)') %>%
  html_text2() %>%
  as.integer() # Converts to integer

# Uses the previous two functions to scrape all of the information into an object
kp_scraped_data <- read_war_page(kp_war_url, max_page)

# Saves the scraped data as cached file .rds
saveRDS(kp_scraped_data, cache_file)
print(paste0('Data Succesfully Compiled')) # Message if succesfully compiled

return(kp_scraped_data) # Returning the data
}



kp_data <- kp_scraper()

```

```{r Getting Guardian Data with API}

# Checks for cached file and if it exists already, it uses that. Since the file is fixed in time, it doesn't check for age. If file doesn't exist yet, it queries the API for all content on ukraine and war from the start of invasion to exactly 4 years later
get_guardian_content <- function(cache_file = 'gu_ua_war_content.rds', force_refresh = FALSE) {
  if (file.exists(cache_file) && !force_refresh) {
    print('Loading cached data') 
    return(readRDS(cache_file))  }
    gu_ua_war_content <- gu_content(query = 'ukraine AND war',
                                from_date = '2022-02-24',
                                to_date = '2026-07-28')
saveRDS(gu_ua_war_content, cache_file)
  gu_ua_war_content
}

# Querying ukraine AND war for all the news with ukraine and war in id
gu_ua_war_content <- get_guardian_content()



# Both last scraped the 28th of July 2026

```


```{r Kyiv Post Preprocessing and Cleaning}


# Kyiv Post preprocessing and minimal cleaning of body text
kp_articles <- kp_data %>%
  filter(str_detect(link, '/post/\\d+')) %>%
  mutate(
    pub_date_raw = str_extract(date_uncleaned, '\\b[A-Z][a-z]{2,}\\.? \\d{1,2}, \\d{4}\\b' # Checks for the date in Kyiv Post format 'July 20, 2023'
), 
    pub_date = mdy(pub_date_raw),  # lubridate function that turns the m-d-year character format into a date format
    pub_month = floor_date(pub_date, 'month'), # extracts the floor date per month
    pub_year = as.integer(year(pub_date)), # extracts the year
    month_of_year = month(pub_date, label = T), # extractws the month, without any other information
    years_passed = as.integer(floor(time_length(interval(as.Date('2022-02-23'), pub_date), 'years'))), # calculates years passed
    war_year = paste0('Year of war: ',years_passed + 1), # creates a 'years of war' variable that measures the years the war is ongoing based on the date of publication via years_passed
) %>%
  mutate(body_text = str_replace_all(body_text_uncleaned, '\\R+', ' '), # basic cleaning of body text by removing all unneccesary html labels
        body_text = str_replace_all(body_text, regex('JOIN US ON TELEGRAM\\s*Follow our coverage of the war on the @\\w+', ignore_case = T), ' '), # remove standard hidden socials KP
        body_text = str_replace_all(body_text, regex('\\badvertisement\\b', ignore_case = T), ' '), # remove standard hidden advertisement
        body_text = str_replace_all(body_text, regex('other topics of interest', ignore_case = T), ' '), # remove the other topics of interest tag
        body_text = str_replace_all(body_text, regex('Follow our coverage of the war on the @Kyivpost_official.', ignore_case = T), ' '), # remove the follow our coverage
        body_text = str_replace_all(body_text, pattern = regex('zelensky', ignore_case = T), replacement = 'zelenskyy'),
        title = str_replace_all(title, pattern = regex('zelensky', ignore_case = T), replacement = 'zelenskyy'), # Equalizing the spelling of zelensky to make comparison possible - universal accepted English spelling 
        body_text = str_squish(body_text)
  ) %>% 
  filter(pub_date >= '2022-06-05', body_text != 'NA', body_text != '', !is.na(body_text)) %>% # Earliest date of news KP after the takeover - important caveat
  add_count(pub_date, name = 'articles_daily') %>%
  add_count(pub_month, name = 'articles_monthly') %>%
  add_count(pub_year, name = 'articles_yearly') %>%
  add_count(war_year, name = 'articles_war_year') %>%
  rowid_to_column(var = 'id') %>%
  arrange(id)


kp_articles

kp_data$body_text_uncleaned[1]
kp_articles$body_text[2]


```

[![Support Ukraine Badge](https://bit.ly/support-ukraine-now)](https://github.com/support-ukraine/support-ukraine)


Mention here that the initial plan was to scrape the Kyiv independent - but that specific war content was only available on the war page, which was a javascript page with no page tags. Instead, I tried to use chromote within Rvest to mimick chrome devtools within a live html read of the war page to then click thousands of times on the load more button. However, after careful consideration this was seen as too cumbersome and prone to crash. A different approach was to look at the xml website page, which had an overview of every link to the posts. The problem was that there was no way to filter for only articles about war here.




Why Ukraine and war? Because Kyiv independent is the main source and they have a dedicated news section for the war, which we use as the benchmark.


```{r Basic exploration of the Guardian Database Structure}

# Check if it starts at starting date 
gu_ua_war_content %>%
  arrange(web_publication_date)

# Which types to use? - articles
unique(gu_ua_war_content$type)

# Which sections we take into account?
unique(gu_ua_war_content$section_id)

gu_ua_war_content %>%
  filter(section_id == 'world')

gu_ua_war_content %>%
  filter(section_id == 'us-news')

gu_ua_war_content %>%
  filter(section_id == 'business')

gu_ua_war_content %>%
  filter(section_id == 'uk-news')

gu_ua_war_content %>%
  filter(section_id == 'books') 

# A lot of the sections are only tangentially related to the russo-ukrainian war. for that reason we decided only to take into account the 'world' section, which is most in line with the war section of the kyiv independent 

gu_articles_raw <- gu_ua_war_content %>%
  filter(section_id == 'world') %>%
  filter(type == 'article')



```


```{r The Guardian Preprocessing and Cleaning}

# Adding dates for days, months, and years. Adding article counts per day, month, and year
gu_articles_full <- gu_articles_raw %>%
  mutate(pub_date = as.Date(web_publication_date),
         pub_month = floor_date(pub_date, 'month'),
         pub_year = year(pub_date),
         month_of_year = month(pub_date, label = T),
         years_passed = as.integer(floor(time_length(interval(as.Date('2022-02-23'), pub_date), 'years'))), # Creates a years_passed variable that measures the floor of the years past as an interval between the start of the war and the publication date
         war_year = paste0('Year of war: ',years_passed + 1)) %>% # creates a 'years of war' variable that measures the years the war is ongoing based on the date of publication via years_passed
  mutate(body_text = str_replace_all(body_text, pattern = regex('zelenskiy', ignore_case = T), replacement = 'zelenskyy'),
        headline = str_replace_all(headline, pattern = regex('zelenskiy', ignore_case = T), replacement = 'zelenskyy')) %>%
          filter(body_text != 'NA', body_text != '', !is.na(body_text)) %>%
  add_count(pub_date, name = 'articles_daily') %>%
  add_count(pub_month, name = 'articles_monthly') %>%
  add_count(pub_year, name = 'articles_yearly') %>%
  add_count(war_year, name = 'articles_war_year') %>%
  rename(old_id = 'id') %>%
  rowid_to_column(var = 'id') %>%
  arrange(id)



gu_articles <- gu_articles_full %>%
  filter(pub_date >= '2022-06-05')




```

```{r Basic Descriptives of Corpi - KP and the Guardian}



kp_body_descriptives <- kp_articles %>%
  mutate(words_per_article = str_count(body_text, pattern = '\\w+')) %>%
  summarize(median = median(words_per_article), average = mean(words_per_article), std = sd(words_per_article), minimum = min(words_per_article), maximum = max(words_per_article), na.rm = T)

kp_body_descriptives


gu_body_descriptives <- gu_articles %>%
  mutate(words_per_article = str_count(body_text, pattern = '\\w+')) %>%
  summarize(median = median(words_per_article), average = mean(words_per_article), std = sd(words_per_article), minimum = min(words_per_article), maximum = max(words_per_article), na.rm = T)

gu_body_descriptives

# From this we can see why there are more tokens in The Guardians corpus despite having fewer articles. The average word count per article is much higher, while the Kyiv Post has a lot more short form articles. However, we also notice that there are a lot of articles with low word count. 

kp_low_word_count <- kp_articles %>%
  mutate(words_per_article = str_count(body_text, pattern = '\\w+')) %>%
  filter(words_per_article < 50)
kp_low_word_count

# From this we garner that all of these are listacles that are not written by the Kyiv Post itself, and rather are taken from the ISW Russian Offensive Campaign Assessment or the British Defence Intelligence. These are not written by any authors of the Kyiv Post itself and are thus taken out of the analysis. 


kp_articles <- kp_articles %>%
  filter(
    !str_detect(title, regex('^ISW Russian Offensive Campaign Assessment.*', ignore_case = T)),
    !str_detect(title, regex('^British Defence Intelligence.*', ignore_case = T)),
    !str_detect(authors, regex('^ISW', ignore_case = T))
  )


# However, we're likely missing more list articles than just these ones. These are taken care of in subsequent runs of scraping by adjusting the scraper. 

```


```{r Kyiv Post hashtag analysis Body Text}

# Are hash-tags meaningful to Kyiv Post war reporting? 
kp_hashtags <- kp_articles %>%
  mutate(
    count_tags = str_count(body_text, '#\\w+'), # regex detects hashtags in the body text
    hash_tags = str_extract_all(body_text, '#\\w+'),
    ) %>%
  select(body_text, count_tags, hash_tags) %>%
  arrange(desc(count_tags))

kp_hashtags %>%
  filter(count_tags > 0)

# 430 articles of 12000+ contain hashtags - which is a meaningful amount


kp_hashtags_content <- unnest(kp_hashtags, cols = c(hash_tags), keep_empty = F)

kp_hashtags_content


```



```{r Guardian Hashtag Analysis Body Text}


# Are hash-tags meaningful to the Guardian war reporting? 
gu_hashtags <- gu_articles %>%
  mutate(
    count_tags = str_count(body_text, '#\\w+'), # regex detects hashtags in the body text
    hash_tags = str_extract_all(body_text, '#\\w+'),
    ) %>%
  select(body_text, count_tags, hash_tags)


# Unnesting the lists to see which hash-tags are used and if they're useful. For this analysis, hash-tags seem not very that relevant, but some words seem 
gu_hashtags_content <- unnest(gu_hashtags, cols = c(hash_tags), keep_empty = T)

# Hashtags ordered by count
gu_hash_count <- gu_hashtags_content %>%
  filter(hash_tags != 'NA') %>%
  count(hash_tags) %>%
  arrange(desc(n)) 

gu_hash_count

# From here we can see that most hashtags are used sparingly in the body text, therefore, the removal of hashtags will do no harm to the analysis. 


# Because of that we decided that we remove all hashtags with words attached to them, because hashtags like #block only contaminate the findings, and we're interested in the content of the article. Hashtags like #Ukraine and #Russia are so underrepresented and are expected to be words that show up in our further analysis - in the content of the article - anyway, that we also delete them 



```




```{r Creating War Anniversaries Tibble for Annotations}

# Creating Events tibble 
war_anniversaries <- tribble(
  ~date, ~label,
  '2023-02-24', 'End of First Year of War',
  '2024-02-24', 'End of Second Year of War',
  '2025-02-24', 'End of Third Year of War',
  '2026-02-24', 'End of Fourth Year of War', 
) %>%
  mutate(date = as.Date(date))


# Approximating the y height on the graph for annotations - The Guardian 
gu_y_height <- approx(x = gu_articles_full$pub_month, 
                       y = gu_articles_full$articles_monthly, 
                       xout = war_anniversaries$date)$y

# Adding the y height to the war_anniversaries tribble for proper annotation on graph - The Guardian
gu_war_anniversaries <- war_anniversaries %>% 
  mutate(y_on_line = gu_y_height)



# Approximating the y height on graph for annotations - KP
kp_y_height <- approx(x = kp_articles$pub_month, 
                       y = kp_articles$articles_monthly, 
                       xout = war_anniversaries$date)$y

# Adding the y height to the war_anniversaries tribble for proper annotation on graph - KP
kp_war_anniversaries <- war_anniversaries %>% 
  mutate(y_on_line = kp_y_height)


```



```{r Guardian Exploratory Visualizations - Number of Articles}

# Exploratory Visualization - Monthly Articles Total Timeline 
gu_articles_plot_total <- ggplot(gu_articles_full, aes(x = pub_month, y = articles_monthly)) + 
  geom_line(color = '#FAF9F6', linewidth = .6, show.legend = FALSE) + 
  geom_vline(data = gu_war_anniversaries, 
            aes(xintercept = date),         
            color = '#FAF9F6',           
            size = .5,
            linetype = 5) +
  geom_text(data = gu_war_anniversaries, 
             aes(x = date, y = y_on_line, label = label), , 
             size = 4, 
             color = '#FAF9F6',
             nudge_x = -25,
             nudge_y = 400,
             angle = 90) +
  geom_point(data = gu_war_anniversaries, 
             aes(x = date, y = y_on_line), 
             color = '#FFD700', size = 1.5) +
  labs(title = 'The Guardian: Amount of Articles per Month',
       x = 'Months', 
       y = 'Articles Count', 
       caption = 'Data from theguardian.com') +
  scale_x_date(date_breaks = '1 month', # Breaks every 2 months
               date_labels = '%b-%Y', # month abbreviated - year format 
               limits = c(as.Date('2022-02-01'), as.Date('2026-08-01')), 
               expand = c(0,0)) + # Removes dead space
  scale_y_continuous(breaks = seq(0, 700, 50),
                     limits = c(0,700)) +
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6, color = '#FAF9F6'),
        axis.text.y = element_text(size = 6, color = '#FAF9F6'),
         panel.background = element_rect(fill = '#0057B7'),
         plot.background = element_rect(fill = '#0057B7'),
         plot.title = element_text(color = '#FAF9F6'),
         axis.title = element_text(color = '#FAF9F6'),
         axis.line = element_line(color = '#FAF9F6'),
         axis.ticks = element_line(color = '#FAF9F6'),
         plot.caption = element_text(color = '#FAF9F6'))




# Clearer view is by segmenting the years into one graph and giving each year a different color

gu_articles_plot_year <- ggplot(gu_articles_full, aes(x = month_of_year, y = articles_monthly, group = as.factor(pub_year), color = as.factor(pub_year))) + 
  geom_line(linewidth = .8, show.legend = T) + 
  labs(title = 'The Guardian: Amount of Articles per Month',
       x = 'Months', 
       y = 'Articles Count', 
       caption = 'Data from theguardian.com') +
  scale_y_continuous(breaks = seq(0, 700, 50),
                     limits = c(0,700)) +
  theme_excel_new() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6, color = '#FAF9F6'),
        axis.text.y = element_text(size = 6, color = '#FAF9F6'),
         panel.background = element_rect(fill = '#0057B7'),
         plot.background = element_rect(fill = '#0057B7'),
         plot.title = element_text(color = '#FAF9F6'),
         axis.title = element_text(color = '#FAF9F6'),
         axis.line = element_line(color = '#FAF9F6'),
         axis.ticks = element_line(color = '#FAF9F6'),
         plot.caption = element_text(color = '#FAF9F6'),
         legend.background = element_rect(fill = '#0057B7'),
         legend.text = element_text(color = '#FAF9F6'))



```

```{r KP Exploratory Visualizations - Number of Articles}

# Exploratory Visualization - Monthly Articles Total Timeline 
kp_articles_plot_total <- ggplot(kp_articles, aes(x = pub_month, y = articles_monthly)) + 
  geom_line(color = '#FAF9F6', linewidth = .6, show.legend = FALSE) + 
  geom_vline(data = kp_war_anniversaries, 
            aes(xintercept = date),         
            color = '#FAF9F6',           
            size = .5,
            linetype = 5) +
  geom_text(data = kp_war_anniversaries, 
             aes(x = date, y = y_on_line, label = label), , 
             size = 4, 
             color = '#FAF9F6',
             nudge_x = -25,
             nudge_y = 200,
             angle = 90) +
  geom_point(data = kp_war_anniversaries, 
             aes(x = date, y = y_on_line), 
             color = '#FFD700', size = 1.5) +
  labs(title = 'Kyiv Post: Amount of Articles per Month',
       x = 'Months', 
       y = 'Articles Count', 
       caption = 'Data from Kyivpost.com') +
  scale_x_date(date_breaks = '1 month', # Breaks every 2 months
               date_labels = '%b-%Y', # month abbreviated - year format 
               limits = c(as.Date('2022-02-01'), as.Date('2026-08-01')), 
               expand = c(0,0)) + # Removes dead space
  scale_y_continuous(breaks = seq(0, 700, 50),
                     limits = c(0,700)) +
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6, color = '#FAF9F6'),
        axis.text.y = element_text(size = 6, color = '#FAF9F6'),
         panel.background = element_rect(fill = '#0057B7'),
         plot.background = element_rect(fill = '#0057B7'),
         plot.title = element_text(color = '#FAF9F6'),
         axis.title = element_text(color = '#FAF9F6'),
         axis.line = element_line(color = '#FAF9F6'),
         axis.ticks = element_line(color = '#FAF9F6'),
         plot.caption = element_text(color = '#FAF9F6'))




# Clearer view is by segmenting the years into one graph and giving each year a different color

kp_articles_plot_year <- ggplot(kp_articles, aes(x = month_of_year, y = articles_monthly, group = as.factor(pub_year), color = as.factor(pub_year))) + 
  geom_line(linewidth = .8, show.legend = T) + 
  labs(title = 'Kyiv Post: Amount of Articles per Month',
       x = 'Months', 
       y = 'Articles Count', 
       caption = 'Data from kyivpost.com') +
  scale_y_continuous(breaks = seq(0, 700, 50),
                     limits = c(0,700)) +
  theme_excel_new() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6, color = '#FAF9F6'),
        axis.text.y = element_text(size = 6, color = '#FAF9F6'),
         panel.background = element_rect(fill = '#0057B7'),
         plot.background = element_rect(fill = '#0057B7'),
         plot.title = element_text(color = '#FAF9F6'),
         axis.title = element_text(color = '#FAF9F6'),
         axis.line = element_line(color = '#FAF9F6'),
         axis.ticks = element_line(color = '#FAF9F6'),
         plot.caption = element_text(color = '#FAF9F6'),
         legend.background = element_rect(fill = '#0057B7'),
         legend.text = element_text(color = '#FAF9F6'))


```

```{r KP and Guardian comparison of # articles}

kp_articles_plot_total/gu_articles_plot_total

kp_articles_plot_year/gu_articles_plot_year

```





```{r The Guardian and Kyiv Post Tidytext tokenization}

# Stopwords used 
head(stop_words)



# For comparison between words, we use the corpi of the Guardian and Kyiv Post that starts on the fifth of June, considering it is the first post that is available on the Kyiv Post website after the restructure. 

# Unnesting words as tokens from body_text - The Guardian
gu_body_tokens_raw <- gu_articles %>%
  unnest_tokens(word, body_text) %>% # default is words
  anti_join(stop_words) # anti-join that removes standard stopwords (tidytext package)

# Unnesting words as tokens from headline - The Guardian
gu_headline_tokens_raw <- gu_articles  %>%
  mutate(
    headline = str_remove(headline, regex("War briefing:?\\s*", ignore_case = TRUE)) # Removes "War briefing" (and any trailing colon/spaces) only at the start of a headline 
  ) %>%
  unnest_tokens(word, headline) %>%
  anti_join(stop_words) 


kp_body_tokens_raw <- kp_articles %>%
  unnest_tokens(word, body_text) %>%
  anti_join(stop_words)

kp_title_tokens_raw <- kp_articles %>%
  unnest_tokens(word, title) %>%
  anti_join(stop_words)

kp_lead_tokens_raw <- kp_articles %>%
  unnest_tokens(word, lead_text) %>%
  anti_join(stop_words)



```



```{r The Guardian and KP lemmatization}


kp_body_tokens <- kp_body_tokens_raw %>% 
  mutate(word = lemmatize_words(word)) 


kp_title_tokens <- kp_title_tokens_raw %>% 
  mutate(word = lemmatize_words(word))




gu_body_tokens <- gu_body_tokens_raw %>% 
  mutate(word = lemmatize_words(word))

gu_headline_tokens <- gu_headline_tokens_raw %>% 
  mutate(word = lemmatize_words(word))


```



```{r The Guardian Tokens Frequencies Visualizations}

# Guardian Tokens Body Text Frequencies visualizations - top 10 words across all years
gu_body_tokens_viz <- gu_body_tokens %>%
  count(word, sort = T) %>%
  slice_max(n = 10, order_by = n) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
    geom_col(fill = '#FAF9F6') +
    labs(title = 'The Guardian Body Text Top 10 Words Across All Years', x = 'Count', y = 'Word') +
    theme_excel_new() +
    theme(
    # Turn the entire background blue
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6')
    )



# Guardian Tokens Body Text Frequencies visualizations - top 10 words faceted by year of war
gu_body_tokens_viz_yearly <- gu_body_tokens %>%
  count(war_year, word, sort = T) %>%
  group_by(war_year) %>%
  slice_max(n = 10, order_by = n) %>% 
  ungroup() %>%
  mutate(word = reorder_within(word, n, war_year)) %>%
  ggplot(aes(n, word, fill = as.factor(war_year))) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~war_year, scales = 'free_y') + # free y scales - reordered for each facet. x scales are fixed for comparison
    scale_y_reordered() +
    labs(title = 'The Guardian Body Text Top 10 Words by Year of War', x = 'Count', y = 'Word') +
    theme_excel_new() +
    scale_x_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale())) + 
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6'), # facet wrap text color white
    ) 



# Guardian Tokens Headline Text Frequencies visualizations - top 10 words across all years
gu_headline_tokens_viz <- gu_headline_tokens %>%
  count(word, sort = T) %>%
  slice_max(n = 10, order_by = n) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
    geom_col(fill = '#FAF9F6') +
    labs(title = 'The Guardian Headline Top 10 Words', x = 'Count', y = 'Word') +
    scale_x_continuous(breaks = seq(0,4000, 500)) +
    theme_excel_new() +
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6')
    )

# Possibly remove briefing from headlines because it doesn't really match here. However. It doesn't matter that much 


# Guardian Tokens Headline Text Frequencies visualizations - top 10 words faceted by year of war
gu_headline_tokens_viz_yearly <- gu_headline_tokens %>%
  count(war_year, word, sort = T) %>%
  group_by(war_year) %>%
  slice_max(n = 10, order_by = n) %>% 
  ungroup() %>%
  mutate(word = reorder_within(word, n, war_year)) %>%
  ggplot(aes(n, word, fill = as.factor(war_year))) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~war_year, scales = 'free_y') + # free y scales - reordered for each facet. x scales are fixed for comparison
    scale_y_reordered() +
    labs(title = 'The Guardian Headlines Top 10 Words by Year of War', x = 'Count', y = 'Word') +
    theme_excel_new() +
    scale_x_continuous(breaks = seq(0, 1500, 500), limits = c(0, 1500)) + 
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6'), # facet wrap text color white
    ) 

# Left briefing in even though it is a title, because it tells us something about the populatiry of the category of such articles and its popularity over the years. 

```


```{r Kyiv Post Tokens Frequencies Visualizations}

# Kyiv Post Tokens Body Text Frequencies visualizations - top 10 words across all years
kp_body_tokens_viz <- kp_body_tokens %>%
  count(word, sort = T) %>%
  slice_max(n = 10, order_by = n) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
    geom_col(fill = '#FAF9F6') +
    labs(title = 'Kyiv Post Body Text Top 10 Words Across All Years', 
    x = 'Count', 
    y = 'Word'
    ) +
    theme_excel_new() +
    theme(
    # Turn the entire background blue
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6')
    )



# Kyiv Post Tokens Body Text Frequencies visualizations - top 10 words faceted by year of war
kp_body_tokens_viz_yearly <- kp_body_tokens %>%
  count(war_year, word, sort = T) %>%
  group_by(war_year) %>%
  slice_max(n = 10, order_by = n) %>% 
  ungroup() %>%
  mutate(word = reorder_within(word, n, war_year)) %>%
  ggplot(aes(n, word, fill = as.factor(war_year))) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~war_year, scales = 'free_y') + # free y scales - reordered for each facet. x scales are fixed for comparison
    scale_y_reordered() +
    labs(title = 'Kyiv Post Body Text Top 10 Words by Year of War', x = 'Count', y = 'Word') +
    theme_excel_new() +
    scale_x_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale())) + 
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6'), # facet wrap text color white
    ) 



# Kyiv Post Tokens Headline Text Frequencies visualizations - top 10 words across all years
kp_title_tokens_viz <- kp_title_tokens %>%
  count(word, sort = T) %>%
  slice_max(n = 10, order_by = n) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
    geom_col(fill = '#FAF9F6') +
    labs(title = 'Kyiv Post Headline Top 10 Words', x = 'Count', y = 'Word') +
    scale_x_continuous(breaks = seq(0,4000, 500)) +
    theme_excel_new() +
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6')
    )



# Kyiv Post Tokens Headline Text Frequencies visualizations - top 10 words faceted by year of war
kp_title_tokens_viz_yearly <- kp_title_tokens %>%
  count(war_year, word, sort = T) %>%
  group_by(war_year) %>%
  slice_max(n = 10, order_by = n) %>% 
  ungroup() %>%
  mutate(word = reorder_within(word, n, war_year)) %>%
  ggplot(aes(n, word, fill = as.factor(war_year))) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~war_year, scales = 'free_y') + # free y scales - reordered for each facet. x scales are fixed for comparison
    scale_y_reordered() +
    labs(title = 'Kyiv Post Headlines Top 10 Words by Year of War', x = 'Count', y = 'Word') +
    theme_excel_new() +
    scale_x_continuous(breaks = seq(0, 1500, 500), limits = c(0, 1500)) + 
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6'), # facet wrap text color white
    ) 


```

```{r KP and The Guardian Word Frequencies Comparisons}

# comparison between total body token frequencies 
kp_body_tokens_viz/gu_body_tokens_viz

# comparison between yearly body token frequencies
kp_body_tokens_viz_yearly/gu_body_tokens_viz_yearly

# comparison between total headline token frequencies
kp_title_tokens_viz/gu_headline_tokens_viz

# comparison between yearly headline token frequencies
kp_title_tokens_viz_yearly/gu_headline_tokens_viz_yearly

```


```{r Frequency comparisons - Kyiv Post and The Guardian}

frequency_body <- bind_rows(
  kp_body_tokens %>% mutate(journal = 'kyiv_post'),
  gu_body_tokens %>% mutate(journal = 'the_guardian')) %>% # -id is to avoid conflict between ids because of class mismatch 
  mutate(word = str_extract(word, "[a-z']+")) %>% # only words, because numbers would be too unique
  count(journal, word) %>%
  group_by(journal) %>%
  mutate(proportion = n/sum(n)) %>%
  select(-n) %>%
  pivot_wider(names_from = journal, values_from = proportion) 



frequency_body %>%
  ggplot(aes(x = kyiv_post, y = the_guardian, color = kyiv_post - the_guardian)) +
  geom_abline(color = '#FAF9F6', linetype = 2) +
  geom_jitter(alpha = .1, size = 2.5, width = .3, height = .3, show.legend = F) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5, show.legend = F) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient2(
    low = '#FF8C00',      # Color for words predominant in The Guardian (negative)
    mid = '#FAF9F6',      # Neutral color for words on the diagonal line (zero difference)
    high = '#FFD700',     # Color for words predominant in Kyiv Post (positive)
    midpoint = 0,
    limits = c(-0.0005, 0.0005),
    oob = scales::squish # Prevents points outside limits from turning gray/invisible
  ) +
  labs(title = 'Body Text Comparison', y = 'The Guardian', x = 'The Kyiv Post') +
  theme_excel_new() +
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6'),
    legend.position = 'none'
    ) 

cor.test(frequency_body$kyiv_post, frequency_body$the_guardian)
cor.test(frequency_body$kyiv_post, frequency_body$the_guardian, method = 'spearman') # spearman test for robustness






frequency_headline <- bind_rows(
  kp_title_tokens  %>% mutate(journal = 'kyiv_post'),
  gu_headline_tokens  %>% mutate(journal = 'the_guardian')) %>% # -id is to avoid conflict between ids because of class mismatch 
  mutate(word = str_extract(word, "[a-z']+")) %>% # only words, because numbers would be too unique
  count(journal, word) %>%
  group_by(journal) %>%
  mutate(proportion = n/sum(n)) %>%
  select(-n) %>%
  pivot_wider(names_from = journal, values_from = proportion) 



frequency_headline %>%
  ggplot(aes(x = kyiv_post, y = the_guardian, color = kyiv_post - the_guardian)) +
  geom_abline(color = '#FAF9F6', linetype = 2) +
  geom_jitter(alpha = .1, size = 2.5, width = .3, height = .3, show.legend = F) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5, show.legend = F) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient2(
    low = '#FF8C00',      # Color for words predominant in The Guardian (negative)
    mid = '#FAF9F6',      # Neutral color for words on the diagonal line (zero difference)
    high = '#FFD700',     # Color for words predominant in Kyiv Post (positive)
    midpoint = 0,
    limits = c(-0.0005, 0.0005),
    oob = scales::squish # Prevents points outside limits from turning gray/invisible
  ) +
  labs(title = 'Headline Comparison', y = 'The Guardian', x = 'The Kyiv Post') +
  theme_excel_new() +
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6'),
    legend.position = 'none'
    ) 


cor.test(frequency_headline$kyiv_post, frequency_headline$the_guardian)
cor.test(frequency_headline$kyiv_post, frequency_headline$the_guardian, method = 'spearman')



```



The words that show up the most in the headlines are expected - Ukraine, Russia, Zelenskyy, Putin, war. But there are some key takeaways here. In year 1 and 2 of the war, invasion is a key word in the vernacular of the Guardian, while in year 3 and 4 invasion doesn't get used as much anymore. This indicates the change in the narrative of the war, from invasion to full scale warfare. Interestingly, Trump shows up in year 3, 4, and 5 as a top 10 word, when he got re-elected as president of the US. Biden, on the other end, didn't make the cut of top 10 words. 

The body texts sees similar results when it comes to words like Ukraine, Russia, Zelenskyy, Putin, and war, and the rise of Trump's name overtime. However, invasion doesn't get mentioned as much, relative to the headlines. Words like forces, military, and people are more common in the body text, which highlights the key actors of the war on the ground. 



Lemmatization - vital for this analysis because of the different forms of words that are used in war reporting. With lemmatization, we keep adjectives like ukrainian and nouns like ukraine separately, which can mean different things in the context of war reporting

Tf-idf weighting is used 

Zelenskyy and Zelenskiy are uniformly replaced by the spelling Zelenskyy.  


US and us? - US is the country, us is the pronoun. In the context of war reporting, both can be relevant. For that reason, we keep them both and don't replace us with US or vice versa -- we need to do this later 

bigrams make sense here to check for ukrainian, russian etc. same with trigrams for body text 


```{r The Guardian bigrams}

gu_body_bigrams <- gu_articles %>%
  unnest_tokens(bigram, body_text, token = 'ngrams', n = 2) %>%
  filter(!is.na(bigram)) # filter out all missings

gu_body_bigrams %>%
  count(bigram, sort = T)


# Bigrams have an obvious amount of stopwords. Let's study that more closely
gu_body_bigrams_sep <- gu_body_bigrams %>%
  separate(bigram, c('word1', 'word2'), sep = ' ')

gu_body_bigrams_filtered <- gu_body_bigrams_sep %>%
  filter(
    !word1 %in% stop_words$word,
    !word2 %in% stop_words$word
    )

gu_body_bigrams_count <- gu_body_bigrams_filtered %>%
  count(word1, word2, sort = T)

gu_body_bigrams_count

gu_body_bigrams_united <- gu_body_bigrams_filtered %>%
  unite(bigram, word1, word2, sep = ' ')



# Headline bigrams 
gu_headlines_bigrams_filtered <- gu_articles %>%
  unnest_tokens(bigram, headline, token = 'ngrams', n = 2) %>%
  filter(!is.na(bigram)) %>% # filter out all missings
  separate(bigram, c('word1', 'word2'), sep = ' ') %>%
  filter(
    !word1 %in% stop_words$word,
    !word2 %in% stop_words$word
    )


gu_headlines_bigram_count <- gu_headlines_bigrams_filtered %>%
  count(word1, word2, sort = T)

gu_headlines_bigram_count


gu_headlines_bigrams_united <- gu_headlines_bigrams_filtered %>%
  unite(bigram, word1, word2, sep = ' ')




```

```{r Kyiv Post bigrams}

kp_body_bigrams <- kp_articles %>%
  unnest_tokens(bigram, body_text, token = 'ngrams', n = 2) %>%
  filter(!is.na(bigram)) # filter out all missings

kp_body_bigrams %>%
  count(bigram, sort = T)


# Bigrams have an obvious amount of stopwords. Let's study that more closely
kp_body_bigrams_sep <- kp_body_bigrams %>%
  separate(bigram, c('word1', 'word2'), sep = ' ')

kp_body_bigrams_filtered <- kp_body_bigrams_sep %>%
  filter(
    !word1 %in% stop_words$word,
    !word2 %in% stop_words$word
    )

kp_body_bigrams_count <- kp_body_bigrams_filtered %>%
  count(word1, word2, sort = T)

kp_body_bigrams_count

kp_body_bigrams_united <- kp_body_bigrams_filtered %>%
  unite(bigram, word1, word2, sep = ' ')



# Title bigrams 
kp_titles_bigrams_filtered <- kp_articles %>%
  unnest_tokens(bigram, title, token = 'ngrams', n = 2) %>%
  filter(!is.na(bigram)) %>% # filter out all missings
  separate(bigram, c('word1', 'word2'), sep = ' ') %>%
  filter(
    !word1 %in% stop_words$word,
    !word2 %in% stop_words$word
    )

kp_titles_bigram_count <- kp_titles_bigrams_filtered %>%
  count(word1, word2, sort = T)

kp_titles_bigram_count


kp_titles_bigrams_united <- kp_titles_bigrams_filtered %>%
  unite(bigram, word1, word2, sep = ' ')




```


```{r Body Bigrams Sentiment Contribution - The Guardian VS Kyiv Post}


gu_negation_words <- gu_body_bigrams_sep %>%
  filter(word1 %in% c('no', 'never', 'nor', 'not', 'without')) %>%
  inner_join(sentiments, by = c(word2 = 'word')) %>%
  count(word1, word2, sentiment, sort = T)

kp_negation_words <- kp_body_bigrams_sep %>%
  filter(word1 %in% c('no', 'never', 'nor', 'not', 'without')) %>%
  inner_join(sentiments, by = c(word2 = 'word')) %>%
  count(word1, word2, sentiment, sort = T)



negation_combined <- bind_rows(
  gu_body_bigrams_sep %>% mutate(journal = 'kyiv_post'),
  kp_body_bigrams_sep %>% mutate(journal = 'the_guardian')) %>% 
  mutate(word1 = str_extract(word1, "[a-z']+"),
        (word2 = str_extract(word1, "[a-z']+")
        )) %>% 
          filter(word1 %in% c('no', 'never', 'nor', 'not', 'without')) %>%
          inner_join(sentiments, by = c(word2 = 'word')) %>%
          count(journal, word1, word2, sentiment, sort= T)

negation_combined



# A complex combined comparison graph of negation bigrams
negation_combined %>%
  mutate(sentiment = ifelse(sentiment == 'positive', 1, -1)) %>%
  mutate(contribution = n * sentiment) %>%
   mutate(
    fill_group = paste(
      ifelse(journal == 'kyiv_post', 'kyiv_post', 'the_guardian'),
      ifelse(contribution > 0, 'positive', 'negative'),
      sep = ' - '
    )) %>%
  arrange(desc(abs(contribution))) %>%
  group_by(word1, journal) %>%
  slice_max(order_by = abs(contribution), n = 10, with_ties = F) %>%
  ungroup() %>%
  ggplot(aes(x = contribution, y = reorder_within(word2, contribution, word1), fill = fill_group, group = journal)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7, show.legend = FALSE) +
  geom_vline(xintercept = 0, linetype = 3, linewidth = .6, color = '#FAF9F6') +
  scale_fill_manual(
    values = c('kyiv_post - positive' = '#FFD700', 'kyiv_post - negative' = '#f75f2c', 
    'the_guardian - positive' = '#5089fb', 'the_guardian - negative' = '#8b0bfb') 
  ) +
  facet_wrap( ~ word1, scales = 'free') +
  labs(x = 'Contribution: Sentiment value * number of occurrences',
       y = 'Words preceded by a negation',
       title = 'Top 10 Words in Sentiment Contribution Preceded by Negation Words Compared') +
        scale_y_reordered() +
         theme_excel_new() +
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    strip.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6')
    ) 
  




```

Checking for negatives - a more comprehensive look at words preceded by 'not'





```{r  Sentiment Analysis}

# This sentiment analysis takes into account negation words if they're in the first part of the bigram 


gu_body_sentiment <- gu_body_bigrams_sep %>%
  inner_join(sentiments, by = c('word2' = 'word')) %>%
  mutate(raw_score = ifelse(sentiment == 'positive', 1, -1)) %>%
  mutate(adjusted_score= ifelse(word1 %in% c('never', 'no', 'nor', 'not', 'without'), -1 * raw_score, raw_score)) %>%
  mutate(adjusted_sentiment = ifelse(adjusted_score == 1, 'positive', 'negative')) %>%
  count(war_year, pub_date, pub_year, pub_month, adjusted_sentiment) %>%
  pivot_wider(names_from = adjusted_sentiment, values_from = n, values_fill = 0) %>%
  mutate(net_sentiment = positive - negative,
  total_sentiment = positive + negative, # total mentions
  ratio_sentiment = net_sentiment / total_sentiment) # density 
  


kp_body_sentiment <- kp_body_bigrams_sep %>%
  inner_join(sentiments, by = c('word2' = 'word')) %>%
  mutate(raw_score = ifelse(sentiment == 'positive', 1, -1)) %>%
  mutate(adjusted_score= ifelse(word1 %in% c('never', 'no', 'nor', 'not', 'without'), -1 * raw_score, raw_score)) %>%
  mutate(adjusted_sentiment = ifelse(adjusted_score == 1, 'positive', 'negative')) %>%
  count(war_year, pub_date, pub_year, pub_month, adjusted_sentiment) %>%
  pivot_wider(names_from = adjusted_sentiment, values_from = n, values_fill = 0) %>%
  mutate(net_sentiment = positive - negative,
  total_sentiment = positive + negative, # total mentions
  ratio_sentiment = net_sentiment / total_sentiment) # density 
  



gu_body_sentiment_viz <- ggplot(gu_body_sentiment, aes(x = pub_date, y = ratio_sentiment, fill = as.factor(war_year))) +
  geom_col(show.legend = F) +
  geom_smooth(show.legend = F, color = '#FAF9F6') +
  facet_wrap(~war_year, scales ='free_x') +
  labs(title = 'The Guardian Body Text Sentiment Negations - by Year of War', x = 'Day of Publication', y = 'Sentiment') +
    scale_x_date(date_breaks = '2 months', # Breaks every 2 months
               date_labels = '%b', # month abbreviated   
               expand = c(0,0)) +
    theme_excel_new() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6, color = '#FAF9F6'),
    panel.background = element_rect(fill = '#0057B7'),
    strip.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6', size = 12),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6')
    ) +
    scale_fill_viridis_d(option = 'turbo')



kp_body_sentiment_viz <- ggplot(kp_body_sentiment, aes(x = pub_date, y = ratio_sentiment, fill = as.factor(war_year))) +
  geom_col(show.legend = F) +
  geom_smooth(show.legend = F, color = '#FAF9F6') +
  facet_wrap(~war_year, scales ='free_x') +
  labs(title = 'Kyiv Post Body Text Sentiment Negations - by Year of War', x = 'Day of Publication', y = 'Sentiment') +
    scale_x_date(date_breaks = '2 months', # Breaks every 2 months
               date_labels = '%b', # month abbreviated   
               expand = c(0,0)) +
    theme_excel_new() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6, color = '#FAF9F6'),
    panel.background = element_rect(fill = '#0057B7'),
    strip.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6', size = 12),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6')
    ) +
    scale_fill_viridis_d(option = 'turbo')








```

```{r Comparisons between all sentiment analyses}

patchwork1 <- (kp_body_sentiment_viz /gu_body_sentiment_viz)

patchwork1 + plot_layout(axis_titles = 'collect', axes = 'collect')


```



```{r The Guardian and Kyiv Post Document-Feature Matrices}

gu_body_tokens_dfm <- quanteda::tokens(gu_articles$body_text,
                         remove_punct = T,
                         remove_symbols = T,
                         remove_numbers = F, #We explicitly keep numbers because they're a key part of war-reporting
                         normalize = T) %>%
  tokens_tolower() %>% 
  quanteda::dfm(verbose = F)




gu_headline_tokens_dfm <- quanteda::tokens(gu_articles$headline,
                         remove_punct = T,
                         remove_symbols = T,
                         remove_numbers = T, 
                         normalize = T) %>%
  tokens_tolower() %>% 
  quanteda::dfm(verbose = F)






kp_body_tokens_dfm <- quanteda::tokens(kp_articles$body_text,
                         remove_punct = T,
                         remove_symbols = T,
                         remove_numbers = F, #We explicitly keep numbers because they're a key part of war-reporting
                         normalize = T) %>%
  tokens_tolower() %>% 
  quanteda::dfm(verbose = F)






kp_title_tokens_dfm <- quanteda::tokens(kp_articles$title,
                         remove_punct = T,
                         remove_symbols = T,
                         remove_numbers = T, 
                         normalize = T) %>%
  tokens_tolower() %>% 
  quanteda::dfm(verbose = F)







gu_body_tokens_dtm <- gu_body_tokens %>%
  cast_dtm(id, word, n)


```



```{r LDA topic analysis}

# Specific stopwords that are used for LDA because they will show up in virtually every article 
domain_stopwords <-  c("ukraine", "russia", "russian", "ukrainian", "ukraine's", "russia's"
    "war", "reporting", "report")



gu_headline_tokens_lda <- LDA(gu_headline_tokens_dfm, k = 2, control = list(seed = 1234))

gu_headline_topics <- tidy(gu_headline_tokens_lda, matrix = "beta")
gu_headline_topics

```

















```{r The Guardian Tokens Frequencies Visualizations}

# Guardian Tokens Body Text Frequencies visualizations - top 10 words across all years
gu_body_tokens_viz <- gu_body_tokens %>%
  count(word, sort = T) %>%
  slice_max(n = 10, order_by = n) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
    geom_col(fill = '#FAF9F6') +
    labs(title = 'The Guardian Body Text Top 10 Words Across All Years', x = 'Count', y = 'Word') +
    theme_excel_new() +
    theme(
    # Turn the entire background blue
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6')
    )



# Guardian Tokens Body Text Frequencies visualizations - top 10 words faceted by year of war
gu_body_tokens_viz_yearly <- gu_body_tokens %>%
  count(war_year, word, sort = T) %>%
  group_by(war_year) %>%
  slice_max(n = 10, order_by = n) %>% 
  ungroup() %>%
  mutate(word = reorder_within(word, n, war_year)) %>%
  ggplot(aes(n, word, fill = as.factor(war_year))) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~war_year, scales = 'free_y') + # free y scales - reordered for each facet. x scales are fixed for comparison
    scale_y_reordered() +
    labs(title = 'The Guardian Body Text Top 10 Words by Year of War', x = 'Count', y = 'Word') +
    theme_excel_new() +
    scale_x_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale())) + 
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6'), # facet wrap text color white
    ) 



# Guardian Tokens Headline Text Frequencies visualizations - top 10 words across all years
gu_headline_tokens_viz <- gu_headline_tokens %>%
  count(word, sort = T) %>%
  slice_max(n = 10, order_by = n) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
    geom_col(fill = '#FAF9F6') +
    labs(title = 'The Guardian Headline Top 10 Words', x = 'Count', y = 'Word') +
    scale_x_continuous(breaks = seq(0,4000, 500)) +
    theme_excel_new() +
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6')
    )

# Possibly remove briefing from headlines because it doesn't really match here. However. It doesn't matter that much 


# Guardian Tokens Headline Text Frequencies visualizations - top 10 words faceted by year of war
gu_headline_tokens_viz_yearly <- gu_headline_tokens %>%
  count(war_year, word, sort = T) %>%
  group_by(war_year) %>%
  slice_max(n = 10, order_by = n) %>% 
  ungroup() %>%
  mutate(word = reorder_within(word, n, war_year)) %>%
  ggplot(aes(n, word, fill = as.factor(war_year))) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~war_year, scales = 'free_y') + # free y scales - reordered for each facet. x scales are fixed for comparison
    scale_y_reordered() +
    labs(title = 'The Guardian Headlines Top 10 Words by Year of War', x = 'Count', y = 'Word') +
    theme_excel_new() +
    scale_x_continuous(breaks = seq(0, 1500, 500), limits = c(0, 1500)) + 
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6'), # facet wrap text color white
    ) 

# Left briefing in even though it is a title, because it tells us something about the populatiry of the category of such articles and its popularity over the years. 

```













WHAT COMES AFTER THIS IS ALL APPENDIX
```{r Kyiv Post Tokens Frequencies Visualizations}

# Guardian Tokens Body Text Frequencies visualizations - top 10 words across all years
kp_body_tokens_viz <- kp_body_tokens %>%
  count(word, sort = T) %>%
  slice_max(n = 10, order_by = n) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
    geom_col(fill = '#FAF9F6') +
    labs(title = 'Kyiv Post Body Text Top 10 Words Across All Years', 
    x = 'Count', 
    y = 'Word'
    ) +
    theme_excel_new() +
    theme(
    # Turn the entire background blue
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6')
    )



# Guardian Tokens Body Text Frequencies visualizations - top 10 words faceted by year of war
kp_body_tokens_viz_yearly <- kp_body_tokens %>%
  count(war_year, word, sort = T) %>%
  group_by(war_year) %>%
  slice_max(n = 10, order_by = n) %>% 
  ungroup() %>%
  mutate(word = reorder_within(word, n, war_year)) %>%
  ggplot(aes(n, word, fill = as.factor(war_year))) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~war_year, scales = 'free_y') + # free y scales - reordered for each facet. x scales are fixed for comparison
    scale_y_reordered() +
    labs(title = 'Kyiv Post Body Text Top 10 Words by Year of War', x = 'Count', y = 'Word') +
    theme_excel_new() +
    scale_x_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale())) + 
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6'), # facet wrap text color white
    ) 



# Guardian Tokens Headline Text Frequencies visualizations - top 10 words across all years
kp_title_tokens_viz <- kp_title_tokens %>%
  count(word, sort = T) %>%
  slice_max(n = 10, order_by = n) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
    geom_col(fill = '#FAF9F6') +
    labs(title = 'Kyiv Post Headline Top 10 Words', x = 'Count', y = 'Word') +
    scale_x_continuous(breaks = seq(0,4000, 500)) +
    theme_excel_new() +
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6')
    )

# Possibly remove briefing from headlines because it doesn't really match here. However. It doesn't matter that much 


# Guardian Tokens Headline Text Frequencies visualizations - top 10 words faceted by year of war
kp_title_tokens_viz_yearly <- kp_title_tokens %>%
  count(war_year, word, sort = T) %>%
  group_by(war_year) %>%
  slice_max(n = 10, order_by = n) %>% 
  ungroup() %>%
  mutate(word = reorder_within(word, n, war_year)) %>%
  ggplot(aes(n, word, fill = as.factor(war_year))) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~war_year, scales = 'free_y') + # free y scales - reordered for each facet. x scales are fixed for comparison
    scale_y_reordered() +
    labs(title = 'Kyiv Post Headlines Top 10 Words by Year of War', x = 'Count', y = 'Word') +
    theme_excel_new() +
    scale_x_continuous(breaks = seq(0, 1500, 500), limits = c(0, 1500)) + 
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6'), # facet wrap text color white
    ) 


```

```{r KP and The Guardian Word Frequencies Comparisons}

# comparison between total body token frequencies 
kp_body_tokens_viz/gu_body_tokens_viz

# comparison between yearly body token frequencies
kp_body_tokens_viz_yearly/gu_body_tokens_viz_yearly

# comparison between total headline token frequencies
kp_title_tokens_viz/gu_headline_tokens_viz

# comparison between yearly headline token frequencies
kp_title_tokens_viz_yearly/gu_headline_tokens_viz_yearly

```


```{r Frequency comparisons - Kyiv Post and The Guardian}

frequency_body <- bind_rows(
  kp_body_tokens %>% mutate(journal = 'kyiv_post'),
  gu_body_tokens %>% mutate(journal = 'the_guardian')) %>% # -id is to avoid conflict between ids because of class mismatch 
  mutate(word = str_extract(word, "[a-z']+")) %>% # only words, because numbers would be too unique
  count(journal, word) %>%
  group_by(journal) %>%
  mutate(proportion = n/sum(n)) %>%
  select(-n) %>%
  pivot_wider(names_from = journal, values_from = proportion) 

frequency_body




# MAKE IT WORK BETTER LATER
frequency_body %>%
  ggplot(aes(x = kyiv_post, y = the_guardian, color = kyiv_post - the_guardian)) +
  geom_abline(color = '#FAF9F6', linetype = 2) +
  geom_jitter(alpha = .1, size = 2.5, width = .3, height = .3, show.legend = F) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5, show.legend = F) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient2(
    low = '#FF8C00',      # Color for words predominant in The Guardian (negative)
    mid = '#FAF9F6',      # Neutral color for words on the diagonal line (zero difference)
    high = '#FFD700',     # Color for words predominant in Kyiv Post (positive)
    midpoint = 0,
    limits = c(-0.0005, 0.0005),
    oob = scales::squish # Prevents points outside limits from turning gray/invisible
  ) +
  labs(y = 'The Guardian', x = 'The Kyiv Post') +
  theme_excel_new() +
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6'),
    legend.position = 'none'
    ) 

cor.test(frequency_body$kyiv_post, frequency_body$the_guardian)
cor.test(frequency_body$kyiv_post, frequency_body$the_guardian, method = 'spearman') # spearman test for robustness






frequency_headline <- bind_rows(
  kp_title_tokens  %>% mutate(journal = 'kyiv_post'),
  gu_headline_tokens  %>% mutate(journal = 'the_guardian')) %>% # -id is to avoid conflict between ids because of class mismatch 
  mutate(word = str_extract(word, "[a-z']+")) %>% # only words, because numbers would be too unique
  count(journal, word) %>%
  group_by(journal) %>%
  mutate(proportion = n/sum(n)) %>%
  select(-n) %>%
  pivot_wider(names_from = journal, values_from = proportion) 

frequency_headline



frequency_headline %>%
  ggplot(aes(x = kyiv_post, y = the_guardian, color = kyiv_post - the_guardian)) +
  geom_abline(color = '#FAF9F6', linetype = 2) +
  geom_jitter(alpha = .1, size = 2.5, width = .3, height = .3, show.legend = F) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5, show.legend = F) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient2(
    low = '#FF8C00',      # Color for words predominant in The Guardian (negative)
    mid = '#FAF9F6',      # Neutral color for words on the diagonal line (zero difference)
    high = '#FFD700',     # Color for words predominant in Kyiv Post (positive)
    midpoint = 0,
    limits = c(-0.0005, 0.0005),
    oob = scales::squish # Prevents points outside limits from turning gray/invisible
  ) +
  labs(y = 'The Guardian', x = 'The Kyiv Post') +
  theme_excel_new() +
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6'),
    legend.position = 'none'
    ) 


cor.test(frequency_headline$kyiv_post, frequency_headline$the_guardian)
cor.test(frequency_headline$kyiv_post, frequency_headline$the_guardian, method = 'spearman')


#### REDO THIS WITH LEMMATIZATION


year_term_counts %>%
  filter(term %in% c("god", "america", "foreign", "union", "constitution", "freedom")) %>%
  ggplot(aes(year, count / year_total)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ term, scales = "free_y") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(y = "% frequency of word in inaugural address")

```



```{r The Guardian dictionary sentiment analysis - by year}


#### FIX DATES LATER

sentiments <- get_sentiments('bing')


# Body_tokens
gu_body_sentiment <- gu_body_tokens %>%
  inner_join(sentiments, by = 'word') %>%
  count(war_year, pub_date, pub_year, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(net_sentiment = positive - negative,
  total_sentiment = positive + negative,
  ratio_sentiment = net_sentiment / total_sentiment) # proportional 


gu_body_sentiment_viz <- ggplot(gu_body_sentiment, aes(x = pub_date, y = ratio_sentiment, fill = as.factor(war_year))) +
  geom_col(show.legend = F) +
  geom_smooth(show.legend = F) +
  facet_wrap(~war_year, scales ='free_x') +
  labs(title = 'The Guardian Body Text Sentiment - Year by Year', x = 'Day of Publication', y = 'Sentiment') +
  scale_x_date(date_breaks = '2 months', # Breaks every 2 months
               date_labels = '%b', # month abbreviated   
               expand = c(0,0)) +
    theme_excel_new() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6, color = '#FAF9F6'),
    panel.background = element_rect(fill = '#0057B7'),
    strip.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6', size = 12),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6')
    ) 



# Headline tokens
gu_headline_sentiment <- gu_headline_tokens %>%
  inner_join(sentiments, by = 'word') %>%
  count(war_year, pub_date, pub_year, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(net_sentiment = positive - negative,
  total_sentiment = positive + negative,
  ratio_sentiment = net_sentiment / total_sentiment)

gu_headline_sentiment_viz <- ggplot(gu_headline_sentiment, aes(pub_date, ratio_sentiment, fill = as.factor(war_year))) +
  geom_col(show.legend = F) +
  geom_smooth(show.legend = F) +
  facet_wrap(~war_year, scales ='free_x') +
  labs(title = 'The Guardian Headline Text Sentiment - Year by Year', x = 'Day of Publication', y = 'Sentiment') +
  scale_x_date(date_breaks = '2 months', # Breaks every 2 months
               date_labels = '%b', # month abbreviated   
               expand = c(0,0)) +
    theme_excel_new() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6, color = '#FAF9F6'),
    panel.background = element_rect(fill = '#0057B7'),
    strip.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6', size = 12),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6')
    ) 




### Equalizing as a ratio? - normalizing? 

```

Headlines seem a lot more balanced when it comes to dictionary style sentiment analysis.

We use the 'bing' dictionary because it is comparitively extensive to the other sentiment dictionaries, which suits the varied corpus of a news outlet with various authors. Comparison: https://medium.com/@laurenflynn1211/comparing-sentiment-analysis-dictionaries-in-r-c695fca64326



```{r Kyiv Post dictionary sentiment - by year}



# Body_tokens
kp_body_sentiment <- kp_body_tokens %>%
  inner_join(sentiments, by = 'word') %>%
  count(war_year, pub_date, pub_year, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(net_sentiment = positive - negative,
  total_sentiment = positive + negative,
  ratio_sentiment = net_sentiment / total_sentiment) # proportional 


kp_body_sentiment_viz <- ggplot(kp_body_sentiment, aes(x = pub_date, y = ratio_sentiment, fill = as.factor(war_year))) +
  geom_col(show.legend = F) +
  geom_smooth(show.legend = F) +
  facet_wrap(~war_year, scales ='free_x') +
  labs(title = 'Kyiv Post Body Text Sentiment - Year by Year', x = 'Day of Publication', y = 'Sentiment') +
  scale_x_date(date_breaks = '2 months', # Breaks every 2 months
               date_labels = '%b', # month abbreviated   
               expand = c(0,0)) +
    theme_excel_new() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6, color = '#FAF9F6'),
    panel.background = element_rect(fill = '#0057B7'),
    strip.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6', size = 12),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6')
    ) 



# Headline tokens
kp_title_sentiment <- kp_title_tokens %>%
  inner_join(sentiments, by = 'word') %>%
  count(war_year, pub_date, pub_year, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(net_sentiment = positive - negative,
  total_sentiment = positive + negative,
  ratio_sentiment = net_sentiment / total_sentiment)

kp_title_sentiment_viz <- ggplot(kp_title_sentiment, aes(pub_date, ratio_sentiment, fill = as.factor(war_year))) +
  geom_col(show.legend = F) +
  geom_smooth(show.legend = F) +
  facet_wrap(~war_year, scales ='free_x') +
  labs(title = 'Kyiv Post Headline Text Sentiment - Year by Year', x = 'Day of Publication', y = 'Sentiment') +
  scale_x_date(date_breaks = '2 months', # Breaks every 2 months
               date_labels = '%b', # month abbreviated   
               expand = c(0,0)) +
    theme_excel_new() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6, color = '#FAF9F6'),
    panel.background = element_rect(fill = '#0057B7'),
    strip.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6', size = 12),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6')
    ) 



```

```{r Sentiment Analysis Comparison between The Guardian and Kyiv Post}

gu_body_sentiment_viz / kp_body_sentiment_viz

gu_headline_sentiment_viz / kp_title_sentiment_viz


```

```{r The Guardian word tf-idf}

gu_body_tfidf <- gu_body_tokens %>%
  count(pub_year, id, word, sort = T) %>%
  bind_tf_idf(term = word, document = id, n = n) 

gu_body_tfidf %>%
  group_by(pub_year, word) %>%
  summarize(tf_idf = max(tf_idf), .groups = 'drop') %>%
  group_by(pub_year) %>%
  slice_max(tf_idf, n = 15, with_ties = F) %>%
  ungroup() %>%
  ggplot(aes(tf_idf, reorder_within(word, tf_idf, pub_year), fill = as.factor(pub_year))) +
  geom_col(show.legend = F) +
  facet_wrap(~pub_year, scales = 'free') +
  labs(x = 'tf-idf score', y = NULL) +
  scale_y_reordered() +
  theme_excel_new() +
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    strip.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6')
    ) 


# tf-idf doesn't deliver much for body_text - why is that???
# The reason for these peculiar results are because short articles where one of these words get mentioned a few times and are almost never mentioned in other articles, results in high tf-idf scores for words that almost never get mentioned. This is somewhat characteristic of the data we're dealing with - huge amount of smaller articles. 

# This is also why it is useless to check this for headlines. However, what can be done is concatonating texts per month or per date


# Tf-idf inherently rewards exclusivity of words. The problem is that a lot of words are going to be exclusive in a varied corpus like news articles 


```


```{r Combined visualization of bigrams that start with negation word}


gu_negation_words %>%
  mutate(sentiment = ifelse(sentiment == 'positive', 1, -1)) %>%
  mutate(contribution = n * sentiment) %>%
  arrange(desc(abs(contribution))) %>%
  group_by(word1) %>%
  slice_max(order_by = abs(contribution), n = 20, with_ties = F) %>%
  ungroup() %>%
  ggplot(aes(x = contribution, y = reorder_within(word2, contribution, word1), fill = contribution > 0)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~word1, scales = 'free') +
  labs(x = 'Contribution: Sentiment value * number of occurrences',
       y = 'Words preceded by a negation',
       title = 'Top 20 Words in Sentiment Contribution Preceded by Negation Words') +
        scale_y_reordered() +
         theme_excel_new() +
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    strip.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6')
    ) 


kp_negation_words %>%
  mutate(sentiment = ifelse(sentiment == 'positive', 1, -1)) %>%
  mutate(contribution = n * sentiment) %>%
  arrange(desc(abs(contribution))) %>%
  group_by(word1) %>%
  slice_max(order_by = abs(contribution), n = 20, with_ties = F) %>%
  ungroup() %>%
  ggplot(aes(x = contribution, y = reorder_within(word2, contribution, word1), fill = contribution > 0)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~word1, scales = 'free') +
  labs(x = 'Contribution: Sentiment value * number of occurrences',
       y = 'Words preceded by a negation',
       title = 'Top 20 Words in Sentiment Contribution Preceded by Negation Words') +
        scale_y_reordered() +
         theme_excel_new() +
    theme(
    panel.background = element_rect(fill = '#0057B7'),
    strip.background = element_rect(fill = '#0057B7'),
    plot.background = element_rect(fill = '#0057B7'),
    plot.title = element_text(color = '#FAF9F6'),
    axis.text = element_text(color = '#FAF9F6'),
    axis.title = element_text(color = '#FAF9F6'),
    strip.text = element_text(color = '#FAF9F6')
    ) 


```